---
title: "Homework3"
output: html_document
---

https://github.com/sydneymbeckett/ENVS-193DS_homework-03.git

```{r}
library(tidyverse)
library(here)
library(gt)
library(janitor)
library(readxl)
ENVS193_FinalData_Sheet1 <- read.csv("ENVS193-FinalData - Sheet1.csv")
ENVS193_FinalData_Sheet1$Steps <- as.numeric(gsub(",", "", ENVS193_FinalData_Sheet1$Steps)) +
#filling the color depending on day of week
```



**Problem 1**
## a.
I could calculate the mean number of steps I took each day to compare average step counts between days when I had class and days when I didn’t, because I think I walk more on days I have class because I walk between classes and from my bike. I could also compare the mean number of steps based on the type of exercise I did, since I expect activities like running to result in more steps than yoga or rest days.

## b.
```{r}
ENVS193_FinalData_Sheet1$Steps <- as.numeric(gsub(",", "", ENVS193_FinalData_Sheet1$Steps))

```


```{r}

ENVS193_FinalData_Sheet1$Day.of.the.Week <- factor(ENVS193_FinalData_Sheet1$Day.of.the.Week,
                                levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) #ordering by days of the week

ggplot(ENVS193_FinalData_Sheet1, aes(x = Day.of.the.Week, #using ggplot with my data and using day of week for x axis
                          y = Steps, #using steps for y axis
                          fill = Day.of.the.Week))+ 
  # Remove commas and convert to numeric
  geom_boxplot(alpha = 0.7, #creating a boxplot and changing transparency using alpha
               
               ) +
  geom_jitter(width = 0.2, size = 1.5, alpha = 0.5, color = "black") + #using jitter to show individual data and changing width size transparency and color of the points
  scale_fill_brewer(palette = "Set2") + #custom color palette
  scale_y_continuous(
    limits = c(0, max(ENVS193_FinalData_Sheet1$Steps, na.rm = TRUE)),
    breaks = scales::pretty_breaks(n = 5),
    labels = scales::comma
  ) +
  labs(title = "Distribution of Steps by Day of the Week", #naming the boxplot
       x = "Day of the Week", #naming x axis
       y = "Number of Steps") + #naming y axis
  
  theme_minimal(base_size = 14) + #using a custom theme
  theme(legend.position = "none", #removing legend becuase it is uneccessary
        axis.text.x = element_text(angle = 45, hjust = 1)) #changes the day of the week to be at an angle so they take up less space
  
```


## c.
Figure 1: Distribution of daily step counts by day of the week. Each box represents the range and median of steps taken, while the individual black points show the underlying data. This visualization highlights how activity levels vary throughout the week, with a higher median on Saturdays, and low variability on Tuesdays. 

## d. 
```{r}
ENVS193_FinalData_Sheet1$Day.of.the.Week <- factor(ENVS193_FinalData_Sheet1$Day.of.the.Week,
                                levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) #ordering by days of the week
summary_table <- ENVS193_FinalData_Sheet1 |> #summarize data using my personal data
  group_by(Day.of.the.Week) |> #grouping the data by day of week
  summarise(`Average Steps` = round(mean(Steps, na.rm = TRUE), 1))  # Round to 1 decimal
summary_table |> 
  gt() |> #using gt package to create a table
  tab_header(
    title = "Average Steps by Day of the Week" #naming the table
  ) |>
  cols_label(
    Day.of.the.Week = "Day", #naming the title of the column on the left day
    `Average Steps` = "Mean Steps" #naming the title of the column on the right mean steps
  )
```



