---
title: "Homework3"
output: html_document
---

https://github.com/sydneymbeckett/ENVS-193DS_homework-03.git

```{r}
library(tidyverse)
library(here)
library(gt)
library(janitor)
library(readxl)
```

**Problem 1**
## a.
I could calculate the mean number of steps I took each day to compare average step counts between days when I had class and days when I didn’t, because I think I walk more on days I have class because I walk between classes and from my bike. I could also compare the mean number of steps based on the type of exercise I did, since I expect activities like running to result in more steps than yoga or rest days.

## b.
```{r}
personal_data <- ENVS193_FinalData_Sheet1 |> #reading in data
  group_by(`Day of the Week`) |> #grouping data by day of week
  summarise(mean_steps = mean(Steps, na.rm = TRUE)) #summarizing the data by taking the mean steps

ENVS193_FinalData_Sheet1$`Day of the Week` <- factor(ENVS193_FinalData_Sheet1$`Day of the Week`,
                                levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) #ordering by days of the week

ggplot(ENVS193_FinalData_Sheet1, aes(x = `Day of the Week`,
                          y = Steps,
                          fill = `Day of the Week`))+ 
  geom_boxplot(alpha = 0.7, 
               outlier.color = "pink",
               outlier.size = 2) +
  geom_jitter(width = 0.2, size = 1.5, alpha = 0.5, color = "black") +
  scale_fill_brewer(palette = "Set2") + #custom color palette
  labs(title = "Distribution of Steps by Day of the Week",
       x = "Day of the Week",
       y = "Number of Steps") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none", #remoxing legend becuase it is uneccessary
        axis.text.x = element_text(angle = 45, hjust = 1))
  
```




